---
title: "MA 331 HW 1"
author: "Aryaman Srivastava"
date: "2023-09-07"
output: pdf_document
pledge: "I pledge my honor that I have abided by the Stevens Honors System"
---

#HW #1 Due 11:59 PM Monday September 11, 2023

All parts of this homework should be submitted to Canvas. Submit the answers to Question 1 and Question 2 parts A and B using handwritten scanned notes or as a latex/typed up pdf file. Submit the answer to Question 2 Part C and Question 3 as the pdf file generated when knitting this report. If need be, combine these pdfs into a single file

##Question 1
Let X be a random variable which takes on the following values:

-1 with probability 1/3
0 with probability 1/3
1 with probability 1/3

Let Y be a random variable that is independent of X but that has the same probability distribution as X

Let Z be a random variable (also independent of X and Y) that takes on the values double that of X, that is, it has the pmf:

-2 with probability 1/3
0 with probability 1/3
2 with probability 1/3


###A) What is the Distribution of X + Y? (Hint: because X and Y are independent, the probability that we draw X to be 0 and Y to be 1 is equal to the product of their individual probabilities, that is, 1/3 * 1/3 = 1/9. In this case, their sum would be 1. However, is that the only way to get a sum of 1?)
-2 with probability 1/9 

-1 with probability 2/9

0 with probability 3/9

1 with probability 2/9

2 with probability 1/9

###B) What is the Distribution of X + Z?
-3 with probability 1/9

-2 with probability 1/9

-1 with probability 2/9

0 with probability 1/9

1 with probability 2/9

2 with probability 1/9

3 with probability 1/9



###C) Find the expected values of X, Y, Z, and X + Y. Show your work
E(X) = -1(1/3) + 0(1/3) + 1(1/3) = 0

E(Y) = -1(1/3) + 0(1/3) + 1(1/3) = 0

E(Z) = -2(1/3) + 0(1/3) + 2(1/3) = 0

E(X+Y) = -2(1/9) + -1(2/9) + 0(3/9) + 1(2/9) + 2(1/9) = 0


###D) Find the variances of X, Y, Z, and X + Y. Show your work
V(X) = E(X^2) - E(X)^2 = 1(1/3) + 0(1/3) + 1(1/3) - 0 = 2/3

V(Y) = E(Y^2) - E(Y)^2 = 1(1/3) + 0(1/3) + 1(1/3) - 0 = 2/3

V(Z) = E(Z^2) - E(Z)^2 = 4(1/3) + 0(1/3) + 4(1/3) - 0 = 8/3

V(X+Y) = E((X+Y)^2) - E(X+Y)^2 = 4(1/9) + 1(2/9) + 0(3/9) + 1(2/9) + 4(1/9) - 0 = 12/9


##Question 2
###A) Find the expected value of a Continuous Random Variable that has the following distribution:

f(x) = 1/2     for 0 <= x <= 2  
        0      elsewhere
        

E(X) = integral of x*f(x) from 0 to 2 = 1/4(2)^2 - 1/4(0)^2 = 1

###B) Find the variance of the same distribution

E(X^2) = integral of x^2*f(x) from 0 to 2 = 1/6(2)^3 - 1/6(0)^2 = 4/3

V(X) = E(X^2) - E(X)^2 = 4/3 - 1 = 1/3


###C) Sample from this distribution 1000 times, and give the empirical mean and variance. You may use the function runif(1000, 0, 2)
        
```{r}
    sampleVector <- runif(1000,0,2)
    mean(sampleVector)
    sd(sampleVector)^2
```


###A) Sample 10,000 samples from an exponential distribution with rate 1, and store the sample in a variable called expData and plot the histogram of these sampled values. You can use the function rexp(). Describe how the histogram looks to you

```{r}
expData <- rexp(10000, 1)
hist(expData)
```
Description: The histogram appears to be skewed right with a majority of the data centered in the interval [0, 2]. If you were to perform the same process multiple times, the distribution would have some variations in the spread and outliers, but the shape and center of the distribution would remain relatively stable. 


###B) For each of 500 replicates, Sample 50 samples from an exponential distribution with rate 1, and store the mean of these samples for each replicate in a vector called expSample50. Plot the histogram of expSample50, and describe how it looks, especially in comparison with the histogram from your sample of size 10,000 from part A

```{r}
expSample50 <- c() 
rep500 <- rep(rexp(50, 1), 500)
for (i in 1:500) {
  expSample50[i] <- mean(rep500[i])
}
hist(expSample50)
```
Description: Similar to the histogram of expData, this histogram seems to be skewed right, with a majority of the data being centered at the interval [0, 1]. As stated previously, the distribution will have some variation if you repeat the process multiple times but the center and shape will remain relatively the same.

###C) For each of 500 replicates, Sample 1000 samples from an exponential distribution with rate 1, and store the mean of these samples for each replicate in a vector called expSample1000. Plot the histogram of expSample100, and describe how it looks, especially in comparison with the histograms from parts A and B. 

```{r}
expSample1000 <- c()
rep500 <- rep(rexp(1000, 1), 500)
for (i in 1:500) {
  expSample1000[i] <- mean(rep500[i])
}
hist(expSample1000)
```
Description: Similar to the last two, this histogram also appears to be right skewed with a majority of the data being at the interval [0, 2] with some variation on the number of outliers and spread if you were to repeat the experiment multiple times.

###D) Compare the means and standard deviations of expData, expSample50, and expSample1000, and add any comments you may have
```{r}
mean(expData)
sd(expData)
mean(expSample50)
sd(expSample50)
mean(expSample1000)
sd(expSample1000)
```
Description: All the means and standard deviations of the three experiments are close to 1, which makes sense as the mean of an exponential distribution should be equal to the rate and the standard deviation 1/rate. There is some slight discrepancy between the means and standard deviations which can be attributed to the sample size/repetition difference between all three experiments. expSample50, which had the least amount of samples is more variable and farther from the expected values for the mean and standard devation. In contrast, expData and expSample1000 is more stable and closer to the expected values.




